Self-interacting dark matter model without dark energy in cosmology

### 2024-12-18 article/article.tex
add cmb reduced paras

\documentclass{aa}
\usepackage{graphicx}
\usepackage{txfonts}
\graphicspath{{./pictures/}}

@@

   \author{Yixuan Zhu}

   \institute{Department of Astronomy, Beijing Normal University.
   Beijing 100875.
   PR China}

   \date{}

   \abstract{}

@@

   \maketitle

@@

\subsection{Model 1: relation to the $\varLambda$CDM model}

@@

   We use the fiducial SN Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances(\cite{Riess_2022}), which gives the $\mu_\text{data}$ and constrains $H_0$
   in advance. To eliminate the influence of $M_b$, we give the $\chi^2$
   as \begin{equation}
      \chi_{\text{SNe}}^2=A-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $A=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})^2/\sigma_i^2,
   B=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})/\sigma_i^2,C=\sum_i^{1701}1/\sigma_i^2$,
   the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.
   \end{equation}
   However, the Eq.(\ref{eq:9}) just need the $H_0$ to caculate the $M_x$
   and we would still use the simple likelihood function as
   \begin{equation}
      \tilde{\chi}_{\text{SNe}}^2=\Delta^{\text{T}}C^{-1}\Delta,
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data.

   $B=\Delta^{\text{T}}C^{-1}$ and $C=\text{sum}(C^{-1})$.

@@

   The quasar gives a higher redshift

@@

   The $\chi^2$ function for the BAO data is defined as
   \begin{equation}
      \chi_{\text{BAO}}^2=\sum_i\frac{(D_{\text{th}}/r_{\text{d}}-D_{\text{data}}/r_{\text{d}})^2}{\sigma_i^2},
   \end{equation}
   where $D$ refers to $D_{\text{M}}$, $D_{\text{H}}$, or $D_{\text{V}}$, which ara given as
   \begin{eqnarray}
      D_{\text{M}}(z)&=&c\int_0^z\frac{\text{d}z'}{H(z')},\\
      D_{\text{H}}(z)&=&\frac{c}{H(z)},\\
      D_{\text{V}}(z)&=&\left[zD_{\text{M}}^2(z)D_{\text{H}}(z)\right]^{1/3},\\
   \end{eqnarray}
   and $r_{\text{d}}$ is the sound horizon at the drag epoch, which is given as
   \begin{equation}
      r_{\text{d}}=\int_{z_{\text{drag}}}^{\infty}\frac{c_s\text{d}z'}{H(z')}.
   \end{equation}

@@

@@

\bibliographystyle{aa}
\bibliography{reference}

>>>>

\documentclass[twocolumn]{aastex631}
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\graphicspath{{./pictures/}}

@@

   \author{Yixuan Zhu}\affiliation{Department of Astronomy, Beijing Normal University.
   Beijing 100875.
   PR China}

   \begin{abstract}

   \end{abstract}

@@

@@

\subsection{Model 1: relation to the LambdaCDM model}

@@

   We use the fiducial SN Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances(\cite{Riess_2022}), which gives the $\mu_{\text{data}}$ and constrains $H_0$
   in advance. To eliminate the influence of $M_b$, we give the $\chi^2$
   as \begin{equation}
      \chi_{\text{SNe}}^2=A-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $A=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})^2/\sigma_i^2,
   B=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})/\sigma_i^2,C=\sum_i^{1701}1/\sigma_i^2$,
   the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.
   \end{equation}
   However, the Eq.(\ref{eq:9}) just need the $H_0$ to caculate the $M_x$
   and we would still use the simple likelihood function as
   \begin{equation}
      \tilde{\chi}_{\text{SNe}}^2=\Delta^{\text{T}}C^{-1}\Delta,
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data.
   $B=\Delta^{\text{T}}C^{-1}$ and $C=\text{sum}(C^{-1})$.

@@

   The quasar gives a higher redshift than SNe Ia.
   We use the QSO dataset from \cite{Lusso_2020}, which gives
   2421 samples with the ultraviolet (UV) and X-ray luminosity.
   The redshift is up to $\simeq7.5$.

   The $L_X-L_{UV}$ relation of quasar is usually written as
   \begin{equation}
      L_X=\beta+\gamma\log L_{UV},
   \end{equation}
   which gives that
   \begin{equation}
      \log_{10}(F_X)=\beta+(\gamma-1)\log_{10}(4\pi)+\gamma\log_{10}(F_{UV})
                    +2(\gamma-1)\log_{10}(d_L),
   \end{equation}
   where $d_L$ is the luminosity distance same as which in the SNe Ia,Eq.15
   so the $\chi^2$ function for the QSO data can be defined as
   \begin{equation}
      \chi_{\text{QSO}}^2=\sum_i^{2421}\frac{(y_{\text{th}}^2-y_{\text{data}}^2)}{s_i^2}
      -\ln(2\pi s_i^2),
   \end{equation}
   where $s_i^2=\mathrm{d}y_i^2+\gamma^2\mathrm{d}x_i^2+\delta^2$ refers to the 
   uncertainties on the $x_i\ (\log_{10}F_X)$ and $y_i\ (\log_{10}F_{UV})$ and
   $\delta_i$ reprensent the instrinsic dispersion.

@@

   The $\chi^2$ function for the BAO data is defined as
   \begin{equation}
      \chi_{\text{BAO}}^2=\sum_i\frac{(D_{\text{th}}/r_{\text{d}}-D_{\text{data}}/r_{\text{d}})^2}{\sigma_i^2},
   \end{equation}
   where $D$ refers to $D_{\text{M}}$, $D_{\text{H}}$, or $D_{\text{V}}$, which ara given as
   \begin{eqnarray}
      D_{\text{M}}(z)&=&c\int_0^z\frac{\text{d}z'}{H(z')},\\
      D_{\text{H}}(z)&=&\frac{c}{H(z)},\\
      D_{\text{V}}(z)&=&\left[zD_{\text{M}}^2(z)D_{\text{H}}(z)\right]^{1/3},
   \end{eqnarray}
   and $r_{\text{d}}$ is the sound horizon at the drag epoch, which is given as
   \begin{equation}
      r_{\text{d}}=\int_{z_{\text{drag}}}^{\infty}\frac{c_s\text{d}z'}{H(z')}.
   \end{equation}

@@

\subsection{Reduced CMB parameters}

   We used the reduced CMB parameters $\{l_A,R,z_{*}\}$ from Planck 2018
   to constrains the model. The reduced CMB parameters are
   \begin{equation}
      l_A=\frac{\pi r(z_{*})}{r_s(z_{*})}
   \end{equation}
   which refer to acoustic scale, and the shift parameter is given as
   \begin{equation}
      R=H_0\sqrt{\Omega_{m0}}r(z_{*})\sim H_0\sqrt{\Omega_{2,0}}r(z_{*}),
   \end{equation}
   where $r(z_{*})$ and $r_s(z_{*})$ are defined as
   \begin{eqnarray}
      r(z_{*})=\int_0^{z_{*}}\frac{z'\mathrm{d}z'}{H(z')},\\
      r_s(z_{*})=\int_{z_{*}}^{\infty}\frac{c_s\mathrm{d}z'}{H(z')}.
   \end{eqnarray}
   As the same as what we do in BAO data, we choose the cross parameter
   $r_s(z_{*})h$ and give the $z_{*}\simeq1090$ to constrain the model.

@@

\bibliography{reference}
\bibliographystyle{aasjournal}
